Clause Size Reduction with \alluip Learning

### 2020-05-10 paper.tex
add cadical results

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\oneuip}{1-UIP\xspace}
\newcommand{\alluip}{all-UIP\xspace}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\stablealluip}{\textit{stable-alluip}\xspace}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
\newcommand{\allUipAct}{\textit{alluip-active}}
\newcommand{\allUipIn}{\textit{alluip-inclusive}}
\newcommand{\allUipEx}{\textit{alluip-exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\config}{\textit{config}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

@@

\section{Implementation and Experiments}
%\subsubsection{Clause Reduction with $\stablealluip$.}
We implemented $\stablealluip$ learning schemes on \text{\MapleBase}
\cite{liang2016maple}, the winner of SAT Race 2015 application
track. We then evaluated these schemes and compare against the \oneuip
baseline on the full set of benchmarks from SAT RACE 2019 main track
which contains 400 instances. We ran our experiments on 2.70 GHz
XeonE5-2680 CPUs, allowing
%\fb{insert time bound}
5000 seconds per instance and a maximum of 12 GB memory.
%\fb{insert RAM bound}.

@@

\begin{figure} 
\begin{center}
\fss{8pt}{10pt}
% \begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  }
\begin{tabular}{|l|l|l|l|} 
\hline
Solver & \#solved (SAT, UNSAT) $\Delta$ & PAR-2 & avg. clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-all-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-all-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-all-act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-all-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-all-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-all-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-all-act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-all-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-all-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-all-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
\hline
$\MapleNineShort$-all-act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-all-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-all-pure & 235 (136, 99) $-$2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-all-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-all-act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

\subsubsection{$\stablealluip$ in Modern SAT solvers.}
\begin{sloppypar}
To validate $\stablealluip$ as a practical learning scheme for modern
SAT solvers, we implemented $\stablealluip$ in the winners of 2017,
2018 and 2019 SAT Race
\cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
and in the $\expSAT$ \cite{MdSolimul2019expMalpe} ($\expSATShort$)
solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
uses random walk simulation to help branching. We chose $\expSATShort$
because (1) it one of the top solvers in the 2019 SAT Race that does
not use chronological backtracking; and (2) the random walk simulation
branching heuristic is different from local branching heuristic (VSIDS
and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
$\allUipEx$. We compared these solver's base \oneuip learning scheme
with $\allUipPure$, $\allUipMin$ and the top two $\stablealluip$
variants, $\allUipAct$ and $\allUipIn$, on the SAT Race 2019 main
track benchmark. We report solved instances, PAR-2 score and the
average clause size.
\end{sloppypar}

@@

On the SAT 2019 race winner $\MapleNineShort$, three out of four
$\stablealluip$ configurations solved instances 6, 5 and 5 more instances
than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
are learned multiple times. We observed that clauses learnt from
$\stablealluip$ configurations are less likely to be duplicated. As an
example, $\allUipMin$ in average, added 12\% less duplicated clauses
into the core clause database than \oneuip. This observation is
surprising, and the cause is unclear.

@@

On $\expSATShort$, three out of four $\stablealluip$ configurations solved
4, 7 and 8 more instances than \oneuip learning. We noticed that
$\allUipAct$ and $\allUipIn$ show significant better performance than
$\allUipMin$ and $\allUipPure$ on this solver. The random walk
simulation branching heuristic, however, didn't impact the performance
of $\stablealluip$ schemes significantly.
%$\stablealluip$ to solve the similar amount of instances with close PAR-2
%scores because the additional random walk exploration allows the
%learning schemes to partially sidestep the activity problem; hence,
%the learning adjustment for variable activities should have less
%impact on the solver's performance. However, we instead, observed that
%both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
%learning scheme. One possible explanation is that $\allUipAct$ and
%$\allUipIn$ schemes could easily over compensate variable activities
%for i-UIP clauses, and $\expSATShort$'s random walk exploration could
%use future search information to mitigate the negative effects of our
%over compensation.

>>>>

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\oneuip}{1-UIP\xspace}
\newcommand{\alluip}{all-UIP\xspace}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\stablealluip}{\textit{stable-alluip}\xspace}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
\newcommand{\allUipAct}{\textit{alluip-active}}
\newcommand{\allUipIn}{\textit{alluip-inclusive}}
\newcommand{\allUipEx}{\textit{alluip-exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\cadical}{\textit{CaDiCaL}}
\newcommand{\defaultcadical}{\textit{CaDiCaL-default}}
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\config}{\textit{config}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

@@

\section{Implementation and Experiments}
%\subsubsection{Clause Reduction with $\stablealluip$.}
We implemented $\stablealluip$ learning schemes on \text{\MapleBase}
\cite{liang2016maple}, the winner of SAT Race 2016 application
track. We then evaluated these schemes and compare against the \oneuip
baseline on the full set of benchmarks from SAT RACE 2019 main track
which contains 400 instances. We ran our experiments on 2.70 GHz
XeonE5-2680 CPUs, allowing
%\fb{insert time bound}
5000 seconds per instance and a maximum of 12 GB memory.
%\fb{insert RAM bound}.

@@

\begin{figure} 
\begin{center}
\fss{8pt}{10pt}
% \begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  }
\begin{tabular}{|l|l|l|l|} 
\hline
Solver & \#solved (SAT, UNSAT) $\Delta$ & PAR-2 & avg. clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-all-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-all-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-all-act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-all-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-all-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-all-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-all-act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-all-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-all-pure & 240 (142, \textbf{98}) +2 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-all-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
\hline
$\MapleNineShort$-all-act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-all-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-all-pure & 235 (136, 99) $-$2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-all-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-all-act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33}
\\
\hline
\hline
 & & &\\
$\defaultcadical$ & 249 (150, 99)  & \textbf{4311.76} & 101.96 \\
\hline
$\defaultcadical$-all-pure & 248 (151, 97) $-$1  & 4373.38 & 82.44  \\
\hline
$\defaultcadical$-all-min & 248 (149, 99) $-$1 & 4398.31 & 46.29 \\ 
\hline
$\defaultcadical$-all-act & \textbf{252} (\textbf{152}, \textbf{100}) +3 & 4331.56 & 47.34 \\
\hline
$\defaultcadical$-all-inclusive & 251 (\textbf{153}, 98) +2 & 4335.88 & \textbf{42.61}
\\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

\subsubsection{$\stablealluip$ in Modern SAT solvers.}
\begin{sloppypar}
To validate $\stablealluip$ as a practical learning scheme for modern
SAT solvers, we implemented $\stablealluip$ in the winners of 2017,
2018 and 2019 SAT Race
\cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
and in the $\expSAT$ \cite{MdSolimul2019expMalpe} ($\expSATShort$)
and $\cadical$~\cite{cadical} solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
uses random walk simulation to help branching. We chose $\expSATShort$
because (1) it one of the top solvers in the 2019 SAT Race that does
not use chronological backtracking; and (2) the random walk simulation
branching heuristic is different from local branching heuristic (VSIDS
and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
$\allUipEx$. We chose $\cadical$ because its default configuration\footnote{$\cadical$'s default configuration is different from the specialized 'sat' configuration, which solved the most SAT instances in the 2019 SAT Race.} (denote as $\defaultcadical$) solved 244 total instances in 2019 SAT Race, the highest among all participants. For this experimented, we used the latest version of $\defaultcadical$ instead of the SAT Race participant~\cite{Armin2019Cadical}.  We compared these solver's base \oneuip learning scheme
with $\allUipPure$, $\allUipMin$ and the top two $\stablealluip$
variants, $\allUipAct$ and $\allUipIn$, on the SAT Race 2019 main
track benchmark. We report solved instances, PAR-2 score and the
average clause size.
\end{sloppypar}

@@

On the SAT 2019 race winner $\MapleNineShort$, all four
$\stablealluip$ configurations solved instances 2, 6, 5 and 5 more instances
than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
are learned multiple times. We observed that clauses learnt from
$\stablealluip$ configurations are less likely to be duplicated. As an
example, $\allUipMin$ in average, added 12\% less duplicated clauses
into the core clause database than \oneuip. This observation is
surprising, and the cause is unclear.

@@

On $\expSATShort$, three out of four $\stablealluip$ configurations solved
4, 7 and 8 more instances than \oneuip learning. We noticed that
$\allUipAct$ and $\allUipIn$ show significant better performance than
$\allUipMin$ and $\allUipPure$ on this solver. The random walk
simulation branching heuristic, however, didn't impact the performance
of $\stablealluip$ schemes significantly.

$\defaultcadical$ with \oneuip solved 249 total instances, 5 more than the result from the SAT 2019 Race.  Applying $\allUipAct$ and $\allUipIn$ further improved the solved instance by 3 an 2, respectively.  The average learnt clause size of \oneuip on $\defaultcadical$ is much larger than other solvers' (101 vs 60) while the average clause size of $\stablealluip$ configurations are similar for all solvers.
